---
title: "Assignment 2"
author: "Edward Kim"
date: "9/20/2020"
output: github_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,message = FALSE, include = FALSE}
library(tidyverse)
library(dplyr)
library(data.table)
library(readr)
library(ggplot2)
```

```{r}
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv", destfile= "individual.csv")
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv", destfile = "regional.csv")

individual <- data.table::fread("individual.csv")
regional <- data.table::fread("regional.csv")

```

# Data Wrangling

## 1. Merge the two data sets 


```{r}

merge <-merge(
  x = individual,
  y = regional,
  by.x = "townname",
  by.y = "townname",
  all.x = TRUE,
  all.y = TRUE)

head(merge)
merge %>% nrow()
```

The two dataset for the individual and regional levels were merged using the location (townname). There is a total of 1200 rows, which confirms that there was no duplicates in the data. There is 49 total variables, which is consistent with the variables present in both datasets. 


## 2. Create new categorical variable 
```{r}
merge <- merge %>% 
  mutate("obesity_level" = case_when( bmi<14 ~ "Underweight",
                                      bmi>=14 & bmi <22 ~"Normal",
                                      bmi>=22 & bmi<24 ~"Overweight",
                                      bmi>=24 ~ "Obese")
  )

merge[,obesity_level:= fifelse(is.na(obesity_level),"Normal",obesity_level)]
summary(merge$bmi)
merge %>% count(obesity_level)
```

The NA's in the data were replaced by the average result for gender and race. 

Because the mean BMI of 18.5 and the median BMI of 17.48 are both in the Healthy category, a majority of the observations would be Healthy.

## 3. Create Categorical Variable "Smoke_gas_exposure" 

```{r}
merge %>% group_by(smoke,gasstove) %>% count()
merge <- merge %>% mutate(smoke_gas_exposure = case_when(smoke==0 & gasstove ==0 ~ "No Exposure",
                                                         smoke==0 & gasstove==1 ~ "Gas Exposure",
                                                         smoke==1 & gasstove ==0~ "Smoke Exposure",
                                                         smoke==1 & gasstove ==1 ~ "Smoke and Gas Exposure",
                                                         )
                          )
merge[,smoke_gas_exposure := fifelse(is.na(smoke_gas_exposure),"Gas Exposure",smoke_gas_exposure)]
merge %>% count(smoke_gas_exposure)
```
Four different categories were created indicating whether the participant was exposed to second-hand smoke, a gas stove, or both.

The NA's in the data were replaced by the average result depending on gender and race. 

## 4. Create four summary tables 
```{r}
merge %>% group_by(townname) %>% 
  summarise(
    fev_avg = mean(fev, na.rm=TRUE),
    asthma.indication = mean(asthma, na.rm = TRUE)
    )

merge <- merge %>% mutate(male.f = case_when( male == 0 ~ "Female",
                                              male == 1 ~ "Male"))

merge %>% group_by(male.f) %>% 
  summarise(
    fev_avg = mean(fev, na.rm=TRUE),
    asthma.indication = mean(asthma, na.rm = TRUE)
    )

merge %>% group_by(obesity_level) %>% 
  summarise(
    fev_avg = mean(fev, na.rm=TRUE),
    asthma.indication = mean(asthma, na.rm = TRUE)
    )

merge %>% group_by(smoke_gas_exposure) %>% 
  summarise(
    fev_avg = mean(fev, na.rm=TRUE),
    asthma.indication = mean(asthma, na.rm = TRUE)
    )
```

# Looking at the Data 


Q1: What is the association between BMI and FEV?
Q2: What is the association between smoke and gas exposure and FEV? 
Q3: What is the assoication between PM2.5 exposure and FEV? 

Initial EDA Checklist:

```{r}
dim(merge)
head(merge)
tail(merge)
summary(merge)
```
Overall, there does not seem to be any issues with the data. We will check the variables that we are analyzing more closely (BMI,FEV,PM2.5, and smoke and gas exposure)

```{r}
summary(merge$fev)
summary(merge$pm25_mass)
summary(merge$bmi)
table(merge$smoke_gas_exposure)
table(merge$obesity_level)
```
The key variables that we are analyzing all seem to have plausible values, and there does not seem to be an error in them. The NA's in the FEV category were all discarded when calculations were run. 

## 1. Facet Plot, BMI vs FEV by "Townname"

```{r, fig.height = 10, fig.width = 12}
merge %>%
  filter(!(townname %in% NA)) %>% 
  ggplot(mapping = aes(x=bmi, y = fev, color = townname))+
  geom_point()+
  labs(title = "Association between BMI and FEV by town")+
  facet_wrap(~townname, nrow = 3)+
  geom_smooth(method = "lm")

```

## 2. Stacked Histograms of FEV by BMI and FEV by smoke/gas exposure

```{r, fig.height =7,fig.width = 9}
merge %>%
  filter(!(townname %in% NA)) %>% 
  ggplot()+
  geom_histogram(mapping = aes(x=fev,fill=obesity_level))+
  scale_fill_brewer(palette="Pastel1")

merge %>% 
  filter(!(townname %in% NA)) %>% 
  ggplot()+
  geom_histogram(mapping = aes(x = fev,fill=smoke_gas_exposure))+
  scale_fill_brewer(palette = "Pastel2")
```